---
title: "HW3 Peer Assessment"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%c %Z')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)

library(car)
```

# Part 1: Logistic Regression
# Background

The objective is to predict whether a passenger survived the Titanic disaster

## Data Description
1. **Survival**: (1 if the passenger survived, 0 if they did not).
2. **Pclass**: Numeric (1,2,3).
3. **Sex**: Categorical (male, female)
4. **Age**: Numeric.
5. **SibSp**: Number of siblings/spouses aboard. 
6. **Parch**: Number of parents/children aboard.
7. **Fare**：Numeric.

## Setup

You can import the data and set up the problem with the following R code:

```{r}
# Import the data
rawdata = read.csv("Titanic.csv", header=TRUE)

# Remove columns that we do not want
rawdata$X <- NULL
rawdata$PassengerId <- NULL
rawdata$Name <- NULL
rawdata$Ticket <- NULL
rawdata$Cabin <- NULL

# Set variables as categorical
rawdata$Pclass<-as.factor(rawdata$Pclass)
rawdata$Sex<-as.factor(rawdata$Sex)
rawdata$SibSp<-as.factor(rawdata$SibSp)
rawdata$Parch<-as.factor(rawdata$Parch)

# Print head of rawdata
head(rawdata)
```

**Note:** For all of the following questions, treat variables **Age** and **Fare** as quantitative variables and **Pclass**, **Sex**, **SibSp**,and **Parch** as categorical variables. Categorical variables have already been converted to factors in the starter code.


# Question 1: Fitting the Model - 9 pts

Use *Survived* as the response variable to fit a logistic regression model using logit as the link function with *Pclass* as the predictor. Call it **model1**.

**(a) 3 pts - Print the model1 summary table. What are the model parameters and estimates?**

```{r}
model1 = glm(Survived ~ Pclass, data = rawdata, family = binomial)
summary(model1)
```
Model parameters are:     
intercept and its estimate is: 0.5432    
pclass2 = -0.6520    
pclass3 = -1.6390    


**(b) 6 pts - Calculate the odds ratio and interpret the estimated coefficient of *Pclass3* with respect to the log-odds of survival and the odds of survival.**

```{r}
exp(coef(model1))
```

The odds for class3 to survive in comparison to class1 is 0.194 times higher. We obtained this by taking the exponential of the log odds which is the coefficient. 0.194 is the odds ratio between class 1 and 3.     
The odds for class2 to survive in comparison to class1 is 0.521 times higher. Again, 0.521 is the odds ratio between class 1 and 2.     
The odds ratio refers to the chance of an event occuring (in our case, survival) in one group to the odds of it occuring in another group.     

The estimated coefficient of Pclass3 (-1.6390) is the log of the odds ratio of the person surviving when his Pclass is 3. 

# Question 2: Statistical Inference - 7 pts 

**(a) 3 pts - Using model1, provide a 95% confidence interval for the coefficient of *Pclass3*.**

```{r}
confint(model1, "Pclass3", level = 0.95)
```

**(b) 4 pts - According to the summary table of model1, is *Pclass3* statistically significant? Is the overall regression statistically significant at the 0.01 significance level?**

```{r}
summary(model1)

stat = model1$null.deviance - model1$deviance
dfs = model1$df.null - model1$df.residual
overall_reg = 1-pchisq(stat, df = dfs)
overall_reg
```
According to the summary of the model, Pclass3 is statistically significant since it has a pvalue of <2e-16, meaning that it is almost zero.    
     
The overall regression is statistically significant (due to a pvalue of 0), which means that the model provides explanatory power.


# Question 3: Goodness of fit - 17 pts

**(a) 5 pts - Assess the goodness-of-fit of model1 using both deviance residuals and Pearson residuals. What do you conclude?**

```{r}
c(deviance(model1), 1-pchisq(deviance(model1),model1$df.residual))
```

```{r}
pearson = residuals(model1, type = "pearson")
pearson.tvalue = sum(pearson^2)
c(pearson.tvalue, 1-pchisq(pearson.tvalue,model1$df.residual))
```
using the deviance residuals, we reject the null hypothesis meaning that the model is not a good fit. However, using the pearson residuals we accept the null hypothesis, meaning that the model IS a good fit.
**(b) 2 pts - Calculate the estimated dispersion parameter for this model. Is this an overdispersed model?**

```{r}
model1_df = model1$df.residual #n-p-1
model1_dev = model1$deviance
overdisp = model1_dev / model1_df
overdisp
```
No, the model isnt overdispersed using a threshold of 2.

**(c) 2 pts - Using *Survived* as the response variable with *Pclass*, *Sex*, *SibSp*, and *Parch* as the predictors and logit as the link function, fit a logistic regression model and call it **model2**.**
```{r}
model2 = glm(Survived ~ Pclass + Sex + SibSp + Parch, family = "binomial", data = rawdata)
```

**(d) 3 pts - Using both deviance residuals and Pearson residuals to assess the goodness-of-fit hypothesis test. What do you conclude? **
```{r}
c(deviance(model2), 1-pchisq(deviance(model2),model2$df.residual))
```

```{r}
pearson2 = residuals(model2, type = "pearson")
pearson2.tvalue = sum(pearson2^2)
c(pearson2.tvalue, 1-pchisq(pearson2.tvalue,model2$df.residual))
```
Both residuals for model2 indicate a good fit. Unlike in model1, where the deviance indicated a not good fit.    

**(e) 3 pts - Calculate the estimated dispersion parameter for this model. Compare model2 with model1 based on the p-values of the deviance residuals and Pearson residuals tests and the dispersion parameters of the two models. Which model is a better fit?**
```{r}
model2_df = model2$df.residual #n-p-1
model2_dev = model2$deviance
overdisp2 = model2_dev / model2_df
overdisp2
```
Model2 seems like a better fit due to having a lower dispersion parameter, and both of its residuals indicating that it is a good fit. Unlike model1 which its deviance suggests that it isnt a good fit.   

**(f) 2 pts - Please provide two methods that can potentially improve the model fit.**
Consider other predictors that can be added to the model that might improve the fit. And, consider transforming the predictors.   

# Question 4: Prediction - 7 pts

Suppose there is a passenger with the following characteristics:

1. **Pclass**: 3

2. **Sex**: male

3. **SibSp**: 2

4. **Parch**: 1

**(a) 4 pts - Predict the passenger's probability of surviving using model1 and model2.**

```{r}
passenger = data.frame(Pclass = 3, Sex = "male", SibSp = 2, Parch = 1)
passenger$Pclass<-as.factor(passenger$Pclass)
passenger$Sex<-as.factor(passenger$Sex)
passenger$SibSp<-as.factor(passenger$SibSp)
passenger$Parch<-as.factor(passenger$Parch)

prediction1 = predict(model1, newdata = passenger, type = "response")
prediction1
```

```{r}
prediction2 = predict(model2, newdata = passenger, type="response")
prediction2
```

Model1 probability of survival = 0.251.

**(b) 3 pts - Compare your predictions. i.e. which model is more reliable based on the analysis?**    
Based on the analysis, model 2 is more reliable.


# Part 2: Poisson Regression
# Background
Bicycle crossing counts for East River bridges crossing into Manhattan, New York City, and we want to focus on Brooklyn Bridge specifically.

## Data Description
1. **High.Temp...F.**: Numeric.
2. **Low.Temp...F.**: Numeric.
3. **Precipitation**: Numeric (range from 0-1).
4. **DayOfWeek**: Categorical (Sun, Mon, Tues, Wed, Thu, Fri, Sat).
5. **Brooklyn.Bridge**: number of bicyclists that cross Brooklyn Bridge in a day.

## Setup

You can import the data and set up the problem with the following R code:

```{r}
# Import data.
casedata <- read.csv("bicycle.csv", header=T)

casedata$Day <- as.Date(casedata$Day)
casedata$Precipitation <- as.numeric(casedata$Precipitation)

casedata$DayOfWeek <- weekdays(casedata$Day)
casedata$DayOfWeek <- as.factor(casedata$DayOfWeek)

casedata_brooklyn <- casedata[,c("High.Temp...F.", "Low.Temp...F.", "Precipitation", "Brooklyn.Bridge", "DayOfWeek")] 

head(casedata_brooklyn)
```

# Question 5: Fitting the Model - 7 pts

Using *Brooklyn.Bridge* as the response variable with *High.Temp...F.*, *Low.Temp...F.*, *Precipitation*, and *DayOfWeek* as the predictors, fit a poisson regression model and call it **poisson_model**.

**(a) 3 pts -  Print the poisson_model summary table. Interpret the estimated coefficient of *Precipitation* with respect to the log expected Brooklyn Bridge count.**
```{r}
poisson_model = glm(Brooklyn.Bridge ~ High.Temp...F. + Low.Temp...F. + Precipitation + DayOfWeek, family = "poisson", data = casedata_brooklyn)

summary(poisson_model)
```
Percipitation has a coefficient of -2.8737164. This means that while holding everything else constant, one unit increase in the percipitation, causes -2.8737164 decrease in the log exprected count.

**(b) 2 pts - Is the overall regression significant?**
```{r}
1-pchisq(poisson_model$null.deviance - poisson_model$deviance, df = poisson_model$df.null-poisson_model$df.residual)
```
The overall regression is statistically significant (due to a pvalue of 0), which means that the model provides explanatory power.

**(c) 2 pts - Perform a goodness-of-fit test on poisson_model using deviance residuals. What do you conclude?**
```{r}
c(deviance(model2), 1-pchisq(deviance(model2),model2$df.residual))
```

# Question 6: Outlier detection - 13 pts

**(a) 2 pts - Create a Q-Q plot of the residuals for the poisson_model.**
```{r}
res <- resid(poisson_model,type="deviance")
qqnorm(res)
```

**(b) 4 pts - Using Cook's Distance with threshold = 1, how many outliers are identified in "poisson_model"? Remove the outliers from the dataset and create poisson_model2 with the new dataset.**
```{r}
cooks_residuals <- cooks.distance(poisson_model, type="deviance")
outliers <- which(cooks_residuals > 1)
length(outliers)
```

```{r}
clean_casedata <- casedata_brooklyn[-outliers,]
poisson_model2 = glm(Brooklyn.Bridge ~ High.Temp...F. + Low.Temp...F. + Precipitation + DayOfWeek, family = "poisson", data = clean_casedata)
```

**(c) 4 pts - Perform the deviance residual test to check the goodness-of-fit of poisson_model2. Is poisson_model2 a good fit? Discuss the appropriateness of using Cook's Distance for poisson regression.**
```{r}
with(poisson_model2, cbind(res.deviance = poisson_model2$deviance, df = poisson_model2$df.residual, p = 1- pchisq(poisson_model2$deviance, poisson_model2$df.residual)))
```
pvalue is zero, meaning that we reject the null hypothesis of good fit, therefore it isnt a good fit.
**(d) 3 pts - Use the alternative method for outlier detection discussed in lecture 13 on the first Poisson model ("poisson_model"). How many outliers are identified? What assumption is made in this approach?**

```{r}
dev_residuals <- residuals(poisson_model, type = "deviance")
outliers_dev <- which(abs(dev_residuals) > qnorm(.99995))
length(outliers_dev)
```
