---
title: "HW3"
date: "2024-01-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(ggplot2)
library(outliers)
usCrime = read.table("uscrime.txt", header = TRUE, stringsAsFactors = FALSE)
crimeCol = usCrime$Crime
crime <- usCrime[,"Crime"]
str(crime)
```

## ISYE 6501 Homework 3      

Question 5.1       
Grubbs test is a method used to detect outliers in a normally distributed data set. By plotting the data in a histogram we can see that the data is somewhat normally distributed and the bloxplot shows us that potential outliers are present.
Grubbs test is predefined by a null hypothesis (there no outliers) and an alternative hypothesis (there is exactly one outlier).               

This first Grubbs test gave a p-value that is considered not significant therefore I removed the data point and tried again. The data point (1969) had a p-value of 0.03 which is considered significant but might be a bit high therefore I removed it and tried again.       

The third try gave a more significant p-value (0.02) which suggests that the data point 1993 is an outlier.            
```{r cars}
plot(crime)
hist(crime)
boxplot(crime) #one sided outlier detection

grubbs.test(crime,type=10)

crime2 <- usCrime[-26,16]
grubbs.test(crime2, type=10) 

crime3 <- usCrime[-4,16]
grubbs.test(crime3, type = 10)

```
Citation:      
https://shorturl.at/fzDQ4      
TA Office Hours     

