---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

uscrime = read.table("uscrime.txt", header = TRUE, stringsAsFactors = FALSE)

```

## ISYE Homework 5 - Linear Regression      

Question 8.1       
I believe a linear regression model would be helpful in looking into and forecasting sales of a store/company. We can add interaction terms for events like discounts/offers/black fridays etc. This way the company could use the model to forecast sales across the seasons and in high/low discounts periods.        
          
Question 8.2         
Linear regression is a way to fit a line to data using least squares and R^2.    
In this homework we are asked to make a linear regression model using US crime data in order to predict the crime level in a hypothetical city, referred to as "point" in the code below. After building the model, the coefficients and intercept of the line produced is the below:         

```{r cars}
model = lm(Crime~., data = uscrime)
model


```

The model gives an R^2 of 0.8 which means that around 80% of the data is taken into consideration while the other 20% is due to randomness or just data that the model doesn't capture. The adjusted R^2 (~0.7) takes into consideration the number of parameters used to fit the line to the data, as not all parameters are of equal significance.         
The R^2 is basically the sum of the distance of the data point from the line squared of all the data points. So a smaller R^2 would mean a better fitted line for the data.      
The p-value reported shows how significant the R^2 (the fit) is. A p-value of 3.59x10^-07 is supposed to be significant since it is less than 0.05.       
However, the output from this model seems to suggest that it overfits the data (ie, being too optimistic).           
          
However, adding the point to the model and having it predict the crime gives:     
```{r}
point = data.frame(M=14.0,So=0,Ed=10.0,Po1=12.0,Po2=15.5,LF=0.640,M.F=94.0,Pop=150,NW=1.1,U1=0.120,U2=3.6,Wealth=3200,Ineq=20.1,Prob=0.04,Time=39.0)


prediction = predict(model, point)
prediction

```

